---
title: "Get language counts from Glottolog for language family and macroarea"
author: "Steven Moran\n"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: github_document
---
  
This script loads the [Glottolog](https://glottolog.org) data and extracts various counts.
  
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
```

Get the Glottolog languoids data.

```{r}
# https://cdstar.shh.mpg.de/bitstreams/EAEA0-D501-DBB8-65C4-0/glottolog_languoid.csv.zip
languoids <- read.csv("glottolog_languoid/languoid.csv", stringsAsFactors = FALSE)
```

How does it look?

```{r}
glimpse(languoids)
```

How many levels are in `level`?

```{r}
table(languoids$level)
```

Unique language family ids. This includes all dialects, empty family_id strings, etc.

```{r}
sort(table(languoids$family_id, exclude = FALSE), decreasing = TRUE)
```

Get just the language level families and their glottocodes.

```{r}
families_glottocodes <- languoids %>% filter(level == "language") %>% filter(family_id != "")
glimpse(families_glottocodes)
```

Get language families and their glottocode counts.

```{r}
families_counts <- families_glottocodes %>% group_by(family_id) %>% summarize(count = n()) %>% arrange(desc(count))
```

There are these many language families in Glottolog -- that are not language isolates.

```{r}
head(families_counts)
dim(families_counts)
```

Get the language isolates.

```{r}
isolates <- languoids %>% filter(family_id == "") %>% filter(parent_id == "") %>% filter(level == "language")
```

There are these many isolates in Glottolog.

```{r}
glimpse(isolates)
```

Let's combine the isolates into the language family counts.

```{r}
isolates_counts <- isolates %>% select(id) %>% rename(family_id = id)
isolates_counts$count <- 1
families_counts <- rbind(families_counts, isolates_counts)
```

There are this many language families, with isolates, reported in Glottolog.

```{r}
glimpse(families_counts)
length(unique(families_counts$family_id))
```


Access the languages and dialects geo data from Glottolog.

```{r, warning=FALSE, message=FALSE}
# languages_and_dialects_geo <- read_csv(url("https://cdstar.shh.mpg.de/bitstreams/EAEA0-D501-DBB8-65C4-0/languages_and_dialects_geo.csv"))
languages_and_dialects_geo <- read_csv("glottolog_languoid/languages_and_dialects_geo.csv")
```

```{r}
glimpse(languages_and_dialects_geo)
```

Get Glottolog macroareas and their language counts.

```{r}
languages_geo <- languages_and_dialects_geo %>% filter(level == "language") %>% filter(macroarea != "") %>% group_by(glottocode)
```

```{r}
head(languages_geo)
```
```{r}
table(languages_geo$macroarea)
```

```{r}
macroarea_counts <- languages_geo %>% group_by(macroarea) %>% summarise(counts = n())
```

```{r}
macroarea_counts
```

Save the counts.

```{r}
save(families_counts, macroarea_counts, file = "glottolog-counts.RData")
```
